Evaluating the Performance and Energy Efficiency of the \textsc{COSMO-ART} Model System

### 2014-05-05 trunk/paper_enahpc14/Sec2:COSMO-ART.tex
cleans up latest version; improves format

The model  system \textsc{Cosmo-art} described  in \citep{Vogel-2009},
is  a  regional to  continental  scale  model  coupled online  to  the
\textsc{Cosmo}   numerical  weather   prediction  and   climate  model
\citep{Baldauf-2011}.   It   incorporates  sophisticated  modules  for
gaseous  chemistry   and  aerosol  dynamics  and   allows  the  online
calculation of  reactive trace  substances and their  interaction with
the  atmosphere.    Detailed  model   description  can  be   found  in
\citep{Bangert-2012, Knote-2011, Knote-2013}.

@@

Establishing effective energy performance benchmarking of a code under
intense development  such as \textsc{Cosmo-art} is  a challenging task
because of  the absolute necessity  that results must  be reproducible
within  an  expected  variance  for  the  duration  of  the  Exa2Green
pro\-je\-ct.  To   define  a  baseline,   it  was  necessary  to   find  a
run-configuration  capable  of   being  recreated  in  all  subsequent
versions.  Here  we give a brief  overview of the model  setup for our
performance and energy efficiency evaluation.

Three-dimensional simulations are performed over large parts of Europe
and the Mediterranean~Sea for April 13$^{th}$ 2010, near the equinox, in order to approximately
have a half  day  of  sun  exposure and  therefore  ensure  a typical
activation of the chemistry cycle.  The calculation domain corresponds
to the CORDEX-EU-44 domain and is covered by a grid of $222\times 216$
points with  a horizontal resolution of  $0.22\,^{\circ}$, i.e., 50~km
in  both directions  and 40~vertical  layers.  These  24-hour forecast
simulations  are not  preceded  by  a spin-up  phase  to build up  the
simulated gaseous and  aerosol concentrations.  This condition means 
that the model initialisation needs a period of adjustment to reach its equilibrium  state and minimise  the effect of
initial conditions for gases and aerosols on the model predictions.

@@

The considered  \textsc{Cosmo-art} model  system is configured  with a
semi-Lagrangian    horizontal   advection    sche\-me    with   tricubic
interpolation  and selective filling  diffusion option  in combination
with   the   dynamical    core   using   Runge-Kutta   time   stepping
\citep{COSMO-PartI-2011}.    It  also   makes  use   of   the  Kinetic
PreProcessor solver (KPP) for  the solution of atmospheric chemistry
ordinary  differential equations \citep{Damian-2002}.   Concerning the
modelling  of  wet  deposition  in  aerosols, the  baseline  has  only
indirect  cloud feedbacks  but  does not  include in-cloud  scavenging
(rainout) and below-cloud  scavenging (washout) yet.  Amongst physical
parameterisations,   precipitation  formation   is   performed  by   a
two-moment cloud microphysics.

>>>>

The  model  system  \cosmoart  described in  \cite{Vogel-2009},  is  a
regional   to  continental   scale   model  coupled   online  to   the
\textsc{Cosmo}   numerical  weather   prediction  and   climate  model
\cite{Baldauf-2011}.    It  incorporates  sophisticated   modules  for
gaseous  chemistry   and  aerosol  dynamics  and   allows  the  online
calculation of  reactive trace  substances and their  interaction with
the  atmosphere.    Detailed  model   description  can  be   found  in
\cite{Bangert-2012, Knote-2011, Knote-2013}.

@@

Establishing effective energy performance benchmarking of a code under
intense development such as \cosmoart is a challenging task because of
the  absolute necessity that  results must  be reproducible  within an
expected variance  for the duration of the  Exa2Green pro\-je\-ct.  To
define  a  baseline, it  was  necessary  to  find a  run-configuration
capable of being recreated in all subsequent versions.  Here we give a
brief  overview of  the model  setup  for our  performance and  energy
efficiency evaluation.

Three-dimensional simulations are performed over large parts of Europe
and the Mediterranean~Sea for  April 13$^{th}$ 2010, near the equinox,
in  order  to  approximately have  a  half  day  of sun  exposure  and
therefore  ensure a typical  activation of  the chemistry  cycle.  The
calculation  domain  corresponds to  the  CORDEX-EU-44  domain and  is
covered  by  a  grid  of  $222\times 216$  points  with  a  horizontal
resolution  of $0.22\,^{\circ}$,  i.e., 50~km  in both  directions and
40~vertical  layers.   These  24-hour  forecast  simulations  are  not
preceded  by a spin-up  phase to  build up  the simulated  gaseous and
aerosol   concentrations.   This  condition   means  that   the  model
initialization needs  a period of adjustment to  reach its equilibrium
state  and minimize  the effect  of initial  conditions for  gases and
aerosols on the model predictions.

@@

The   considered  \cosmoart   model  system   is  configured   with  a
semi-Lagrangian   horizontal    advection   sche\-me   with   tricubic
interpolation  and selective filling  diffusion option  in combination
with   the   dynamical    core   using   Runge-Kutta   time   stepping
\cite{COSMO-PartI-2011}.    It   also  makes   use   of  the   Kinetic
PreProcessor solver  (KPP) for  the solution of  atmospheric chemistry
ordinary  differential equations  \cite{Damian-2002}.   Concerning the
modelling  of  wet  deposition  in  aerosols, the  baseline  has  only
indirect  cloud feedbacks  but  does not  include in-cloud  scavenging
(rainout) and below-cloud  scavenging (washout) yet.  Amongst physical
parameterizations,   precipitation  formation   is   performed  by   a
two-moment cloud microphysics.
